---
title: "Lambda and Transformation Verification"
author: "Liwei Jiang"
execute: 
  echo: false
---

#### Note:

This qmd is dedicated to testing hypotheses of lambda and log transformation.

I copied majority of Kassie's code in file "First Model.qmd". But as we discussed over Wechat, I use source() instead of copying all data wrangling codes. And then I plot and see whether the data needs to be normalized. Then, I check whether lamda needs to be specify or we should leave R to figure out the optimal number of lamda.

#### Loading Packages

```{r}
#| warning: false
#| message: false
library(readxl)
library(tidyverse)
library(lubridate)
library(zoo)
library(here)
library(forecast)
library(knitr)
```

#### Run Data Wrangling

```{r child="Data Wrangling.qmd", eval=FALSE}

# I failed to use source(). But I use child to make sure codes in Data Wrangling.qmd will be executed while rendering (knitting).
# But we still need to execute the codes manually before rendering to have variables and dataframes in the environment. Kind of lame for sure
```

#### Examine the distribution of the original data

```{r}

# create a column in ts_combined as log(load)
ts_combined <- ts_combined[, 1] %>% 
  log() %>% 
  cbind(ts_combined)

# rename columns of ts_combined
colnames(ts_combined) <- c("log(load)", "load", "humidity", "temperature")

# plot log(load) and load to compare
par(mfrow = c(1,2))
plot(ts_combined[, 2], ylab = "Load")
plot(ts_combined[, 1], ylab = "Log(Load)")

# from the plots we can see log transformation doesn't necessarily improve the original data. 
# So, I would recommend we stick with the original data

# drop log(load)
ts_combined <- ts_combined[, -1]

```

##### Test Lambda Issue

```{r}

# use Kassie's auto arima
Model2 <- auto.arima(ts_load,
                     seasonal = FALSE,
                     lambda=0,
                     xreg = fourier(ts_load,
                                    K=c(2,12)
                                    )
                     )
# let R find lambda by itself then compare RMSE
model2_test <- auto.arima(ts_load,
                          seasonal = FALSE,
                          lambda = "auto",
                          xreg = fourier(ts_load,
                                         K=c(2,12)
                                         )
                          )
model2_test1 <- auto.arima(ts_load,
                          seasonal = FALSE,
                          xreg = fourier(ts_load,
                                         K=c(2,12)
                                         )
                          )

accuracy(Model2_for$fitted, ts_load_train) 

accuracy(model2_test$fitted, ts_load_train) 

accuracy(model2_test1$fitted, ts_load_train) 

# without lambda has a lower RMSE, so maybe we don't need to include lambda in our model
```
